# LAGOS Spatial Analysis Part One


```{r LAGOS setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(RApiSerialize) 
library(dplyr)
library(lubridate) #For dealing with date and time
```


```{r LAGOS data-read,  include=FALSE, echo=FALSE}
# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)


# #Lagos download script
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())

#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus

```



```{r, include=FALSE, echo=FALSE}
###Convert to spatial data

#Look at the column names

names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


```{r, include=FALSE, echo=FALSE}
### Subset to only Minnesota

states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



## In-Class work


1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r, echo=TRUE}
iowa <- states %>%
  filter(name == 'Iowa') %>%
  st_transform(2163)

illinois <- states %>%
  filter(name == 'Illinois') %>%
  st_transform(2163)

mapview(iowa) +
  mapview(illinois)

```



2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa combined? How does this compare to Minnesota?

```{r, echo=TRUE}

iowa_lakes <- spatial_lakes[iowa,]
illinois_lakes <- spatial_lakes[illinois,]

```

-Iowa has 4,644 lakes total, and Illinois has 11,822. This combines to about ~16,500+ lakes, which is about half as many as Minnesota contains (29,038). Minnesota has a ton of lakes.


3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r, echo=TRUE, warning=FALSE, message=FALSE}
ggplot(minnesota_lakes, aes(x= lake_area_ha)) + 
  geom_histogram() + 
  scale_x_log10(labels = scales::comma) +
  ylab("Lake Area (ha)") +
  xlab("Number of lakes") +
  ggtitle("Minnesota Lakes")

ggplot(iowa_lakes, aes(x= lake_area_ha)) +
  geom_histogram() +
  scale_x_log10(labels = scales::comma) +
  ylab("Lake Area (ha)") +
  xlab("Number of lakes") +
  ggtitle("Iowa Lakes")

```

-Minnesota has many more lakes that take up over 1,000 hectares, with a few reaching 5,000 hectares. The lakes in Iowa top out around 1,100 hectares.


4) Make an interactive plot of lakes in Iowa and Illinois and color them by lake area in hectares

```{r, echo=TRUE}
Iowanois=rbind(iowa_lakes, illinois_lakes)

Iowanois %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```


5) What other data sources might we use to understand how reservoirs and natural lakes vary in size in these three states? 

-Aquifer depletion/depth, precipitation, long-term climate data, and glacial movement data would help to explain these size differences. Land use change data might also help to explain the manmade lakes. 

